Draft

### 2019-07-24 draft.tex
Coproduct is preserved under nullification - full proof.

\documentclass[12pt]{article}

%\usepackage{answers}

\usepackage{setspace}
%\usepackage{graphicx}
\usepackage{enumitem}
%\usepackage{tikz-cd}
%\usepackage{todonotes}
%\usepackage{multicol}
%\usepackage{mathrsfs}
%\usepackage[margin=1in]{geometry} 
\usepackage{hyperref}% http://ctan.org/pkg/hyperref
\usepackage{cleveref}% http://ctan.org/pkg/cleveref
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{ textcomp }
%\usepackage{ wasysym }
\usepackage{bookmark}
%\usepackage{accents}
\usepackage[parfill]{parskip}

@@

\maketitle
%Below is an example of the problem environment
\section{Type Theory}
\subsection{Introduction}
Type theory is a formal language and deductive system, that is self sufficient in the sense that it need not be formulated as a collection of axioms on top of some other formal system like First Order Logic, instead its deductive system can be built on top of its own formal language. 
\todo{longer sentence to check formating }

@@

\subsection{Type Construction Operations}
Let's have a look at some important type constructions.
\begin{itemize}
	\item Given types $A, B : \mathcal{U}$ we can define the type $A \rightarrow B$ of functions from $A$ to $B$. 
		We can use $\lambda$-abstraction to construct elements of this type. $\lambda x. \Phi$ lies in $A \rightarrow B$ iff for $a : A$ we have $\Phi[a/x] : B$. 
		For $f : A \rightarrow B$ and $a : A$ we have that the application of $f$ on $a$, denoted as $f(a)$ or $f\;a$, lies in $B$.
	\item Given some type $A : \mathcal{U}$ and a family of types $B$ over $A$, $B : A \rightarrow \mathcal{U}$, we have the type of dependent functions $$\prod_{a : A} B(a)$$
	where for $f : \prod_{a : A} B(a)$ and $x : A$ we have $f(x) : B(x)$. 
	As in the case of non-dependent functions, we can use lambda abstraction to construct elements of a dependently-typed function type.
	\item Given $A, B : \mathcal{U}$ we can define the product type $A \times B : \mathcal{U}$. 
	For $a : A$ and $b : B$ we have the pair $(a,b) : A \times B$. 
	We also have the projection functions 
	$$\mathtt{pr}_1 : A \times B \rightarrow A : (a,b) \mapsto a$$
	$$\mathtt{pr}_2 : A \times B \rightarrow B : (a,b) \mapsto b$$
	\item Given $A : \mathcal{U}$ and family of types $B$ over $A$, $B : A \rightarrow \mathcal{U}$, we can define the dependent pair type $$\sum_{a : A} B(a)$$
	Given $x : A$ and $b : B(x)$ we can construct the pair $(x,b) : \sum_{a : A} B(a)$. 
	We have two projection functions, similar to the case of the product type.
	\item Given $A, B : \mathcal{U}$ we can construct the coproduct type $A + B$. 
	We can construct elements of $A + B$ using the functions
	$$\mathtt{inl} : A \rightarrow A+B$$
	$$\mathtt{inr} : B \rightarrow A+B$$
	This induces the induction principle
	$$\mathtt{ind}_{A+B} : \prod_{C : A+B \rightarrow U} \Big(\prod_{a : A}C(\mathtt{inl}\;a)\Big) \rightarrow \Big(\prod_{b : A}C(\mathtt{inr}\;b)\Big) \rightarrow \prod_{x : A+ B}C(x)$$
	\item Given $x, y : A$ we have the \textbf{identity type} $x =_A y$. 
	An element of this type amounts to a proof that $x$ and $y$ are equal. 
	Say $x$ and $y$ are judgmentally equal. 
	This is captured by the element $\mathtt{idp}_x : x =_A y$. 
	The relevant induction principle describes how we can use elements of an identity type
	$$\mathtt{ind}_{=_{A}} : \prod_{C : \prod_{x,y : A} (x =_A y) \rightarrow \mathcal{U}} \Big( \prod_{x : A}C(x,x,\mathtt{idp}_x) \Big) \rightarrow \prod_{x,y : A} \prod_{(p : x =_A y)} C(x,y,p)$$
	The relevant computation gives us the judgmental equality 
	$$\mathtt{ind}_{=_A} (C,c,x,x,\mathtt{idp_x}) \equiv c\; x$$
	We can concatenate those paths whose domains and codomains allow for it. 
	Paths are equivalences. That is if $p : x = y$ is such a  path, we can provide its inverse $p^{-1}$ for which we have in turn a path between $p\cdot p^{-1}$ and $\mathtt{idp}_y$ and another one between $p^{-1}\cdot p$ and $\mathtt{idp}_x$. 
	\item For every type $A$ there is its \textbf{propositional truncation} $\| A \|$. 
	For every element $a : A$ there exists $|a| : \| A\|$. 
	For every $x, y : \| A\|$ we have $x = y$. 
	Given mere proposition $B$ and $f : A \rightarrow B$, the recursion principle gives us $g : \|A\| \rightarrow B$ such that $g(|a|) \equiv f(a)$ for all $a : A$.

@@

\subsection{Important types}
\begin{defn}
We call a type $A$ a \textbf{mere proposition} if for every $a,b : A$ we have $a = b$.
\end{defn}

@@

\subsection{Logic}
Our informal deductions in Type Theory will be reminiscent of First Order Logic ones. 
To be able to use a similar verbiage, we will set down a handful of types, corresponding to the connectives that let us form well-formed formulas in FOL. 
These types need to be mere propositions, so that we can form non-constructive deductions. 
This approach is called `propositions as h-propositions' in the HoTT book. 
In the following, $A$ and $B$ are mere propositions.
\begin{itemize}
\item When we talk of conjunction $A \wedge B$, we mean the product $A \times B$. 
\item We interpret $A \vee B$, as the truncation $\| A + B\|$.
\item We interpret $\forall a\in A,\; P(a)$, where $P(a)$ is a mere proposition for all $a \in A$, as $\prod_{a : A}P(a)$. 
\item We interpret $ \exists a \in A,\; P(a)$, where $P(a)$ is a mere proposition for all $a \in A$, as $\| \Sigma_{a : A} P(a) \|$
\end{itemize}
The use of the above notation is interchangeable in the sections to follow.

\section{Modalities}
%Under the view of types as $(\infty, 1)$-categories, functions between them are functors. 
%We can then view the universe $\mathcal{U}$ itslef as a category. 
\begin{defn}
A \textbf{modality} is a function $\bigcirc : \mathcal{U} \rightarrow \mathcal{U}$ with the following properties.
\begin{enumerate}
	\item For every type $A$ we have a function $\eta_A^\bigcirc : A \rightarrow \bigcirc\; A$
	\item for every $A : \mathcal{U}$ and every type family $B : \bigcirc A \rightarrow \mathcal{U}$ we have a function
	$$\mathtt{ind}_\bigcirc : \Big( \prod_{a : A}\bigcirc (B (\eta_A^\bigcirc\; a)) \Big) \rightarrow \prod_{z : \bigcirc A} \bigcirc (B\; z)$$
	\item For every $f : \prod_{a : A} \bigcirc (B(\eta_A^\bigcirc\; a))$ there is a path $\mathtt{ind}_\bigcirc (f)(\eta_A^\bigcirc\; a) = f\; a$
	\item For all $z,z' : \bigcirc \; A$, the function $\eta_{z=z'}^\bigcirc : (z = z') \rightarrow \bigcirc (z = z')$ is an equivalence.

@@

\section{LLPO}
\begin{defn} 
The Lesser Limited Principle of Omniscience, states that given binary sequence $s : \mathbb{N} \rightarrow \mathbf{2}$ and the fact that there is at most one occurence of $1$ in the sequence, formally 
$$\mathtt{atMost1one} :\prod_{n_1 : \mathbb{N}} \prod_{n_2 : \mathbb{N}} s(n_1) = 1 \rightarrow s(n_2)= 1 \rightarrow n_1 = n_2$$
we can then have by the LLPO a witness for $\mathtt{p_{odd}} \vee \mathtt{p_{even}}$, where $\mathtt{p_{odd}}$ (with $s$ as an implicit argument) is the statement that for all odd positions $n$, $s(n) = 0$, formally $\mathtt{p_{odd}} \equiv \prod_{n : \mathbb{N}} \mathtt{odd}(n) \rightarrow s(n) = 0$. 
Similarly for $\mathtt{p_{even}}$. 
\end{defn}
LLPO can be viewed as a weaker form of the Law of Excluded Middle.

\begin{lem} 
The Law of Excluded Middle implies the Lesser Limited Principle of Omniscience.
\end{lem}

@@

	\item $\neg \mathtt{p_{odd}}$, alongside LEM, implies that there exists odd $n_e : \mathbb{N}$ such that $s(n_e) = 1$. 
	We can now provide $\prod_{n : \mathbb{N}} \mathtt{even}(n) \rightarrow s(n) = 0$. 
	Let $n : \mathbb{N}$. 
	By the definition of $s$, $s(n) = 0 \vee s(n) = 1$.
	We invoke the principle of induction of coproducts and prove LLPO from the disjuncts. 
	\begin{itemize}
		\item $s(n) = 0$, in this case we are done.
		\item $s(n) = 1$. 
		By $\mathtt{atMost1one}$ we have $n = n_e \Rightarrow $ $n$ even and odd which is a contradiction $c : \bot$. 
		Ex Falso,  $\mathtt{efq} : \bot \rightarrow s(n) = 0$. 
		Then $\mathtt{efq}(c) : s(n) = 0$.
	\end{itemize}
\end{itemize}
\end{proof}

@@

\section{Extensional Type Theory}
There exists a model of extensional type theory that validates the following `Independence Principle'
$$\bigg( \prod_{s : \mathbf{2}^\mathbb{N}} P\; s \rightarrow \Big(\prod_{n : \mathbb{N}}s\; n = 0 \Big) \rightarrow \Big\lVert \sum_{z : \mathbb{N}} Q\; s\; z \Big\rVert \bigg)$$
$$\rightarrow \bigg( \prod_{s : \mathbf{2}^\mathbb{N}} P\; s \rightarrow  \Big\lVert \sum_{z : \mathbf{2}} \Big(\prod_{n : \mathbb{N}}s\; n = 0 \Big) \rightarrow Q \; s\; z\Big\rVert \bigg)$$
where $P: \mathbf{2}^\mathbb{N} \rightarrow \mathtt{Prop}$ and $Q : \mathbf{2}^\mathbb{N} \rightarrow \mathbb{N} \rightarrow \mathtt{Prop} $ are families of propositions. 
Putting it plainly, if the left hand side of the above is true, then $z : \mathbb{N}$ does not depend on the proof of the constantness of $s$ to $0$.

@@

In this section we will be working in the context of this model.

Let $A :\equiv \sum_{a: \mathbf{2}^\mathbb{N}}\mathtt{atMost1one}\;a$. 
When referring to elements of $A$ we implicitly mean the first part of the pair. 
Let
$$B : A \rightarrow \mathtt{Prop}$$
$$B :\equiv \lambda\; a.\; \mathtt{p_{odd}}\; a \vee \mathtt{p_{even}}\; a$$
\begin{lem}
$\mathbb{N}$ is $B$-null.
\end{lem}
\begin{proof}
Given $a : A$ and $f : B\; a \rightarrow \mathbb{N}$ we need to prove that there exists unique $f' : \mathbf{1} \rightarrow \mathbb{N}$ through which $f$ factors, in the sense that $f = f' \circ g$, where $g : B\; a \rightarrow \mathbf{1}$ is the sole inhabitant of its function space. 
Functions with $\mathbf{1}$ as their domain are constant. 
We therefore need to find some $z : \mathbb{N}$ so that $\lambda\; \_.\; z$ is $f'$. 
To that end, we will use the following instance of the Independence Principle
$$\bigg( \prod_{s : \mathbf{2}^\mathbb{N}} \Big(\big(\prod_{n : \mathbb{N}} s(n) = a(2 \cdot n)\big) \vee \big(\prod_{n : \mathbb{N}} s(n) = a(2\cdot n +1)\big) \Big) \rightarrow \Big(\prod_{n : \mathbb{N}}s\; n = 0 \Big) \rightarrow \Big\lVert \sum_{z : \mathbb{N}}\prod_{p : B\; a} f\; p = z\Big\rVert \bigg)$$
$$\rightarrow \bigg( \prod_{s : \mathbf{2}^\mathbb{N}} \Big(\big(\prod_{n : \mathbb{N}} s(n) = a(2 \cdot n)\big) \vee \big(\prod_{n : \mathbb{N}} s(n) = a(2\cdot n +1)\big) \Big) \rightarrow  \Big\lVert \sum_{z : \mathbb{N}} \Big(\prod_{n : \mathbb{N}}s\; n = 0 \Big) \rightarrow\prod_{p : B\; a} f\; p = z\Big\rVert \bigg)$$
Note that the left hand side is provable. 
By providing the odd $s_1$ and even $s_2$ subsequences of $a$ to the right hand side, we get hold of $z_1 : \mathbb{N}$ and $z_2 : \mathbb{N} $ respectively. 
Equality on $\mathbb{N}$ is decidable, therefore $z_1 = z_2 \vee z_1 \neq z_2$ is provable. 
Since our goal is a proposition, we can drop the truncation of the disjunction. 
This leaves us with a coproduct.

We will first prove that the existence of a candidate for $f'$ follows from both constituents of the coproduct.
\begin{itemize}
\item First the case where $z_1= z_2$. 
We arbitrarily pick $z_1$ and propose $f' : \equiv \lambda \;\_.\; z_1$. 
By axiom K we have function extensionality, so we reduce proving $f= f' \circ g$ to proving $f\; b = f'\circ g\; b$ for arbitrary $b : B\; a$. \\
$b$ is a truncated coproduct witness. 
As before we can ignore the truncation and take cases on the coproduct. 
\begin{itemize}
\item We first consider the case $\mathtt{p_{odd}}\; a$. 
This tells us that the odd subsequence is constantly $0$, which we can provide to the rhs of the IP alongside with $b$  itself (and $s_1$ as the first argument) to get $f\; = z_1$. We concatenate this with $f' \circ g \; b = z_1$ to conclude this case. 
\item The $\mathtt{p_{even}}$ case is very similar. 
The only difference is that we provide $s_2$ as the first argument to the rhs of the IP and then have $z_1=z_2$ be part of the concatenation of paths when we prove homotopy between $f$ and $f' \circ g$. 
\end{itemize}

>>>>

\documentclass[12pt]{report}

\usepackage{tikz-cd}
%\usepackage{answers}
\usepackage{biblatex}
\addbibresource{bibliography.bib}
\usepackage{setspace}
%\usepackage{graphicx}
\usepackage{enumitem}
%\usepackage{tikz-cd}
%\usepackage{todonotes}
%\usepackage{multicol}
%\usepackage{mathrsfs}
%\usepackage[margin=1in]{geometry} 
\usepackage{hyperref}% http://ctan.org/pkg/hyperref

\usepackage{amsmath}
\usepackage{cleveref}% http://ctan.org/pkg/cleveref
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{ textcomp }
%\usepackage{ wasysym }
\usepackage{bookmark}
%\usepackage{accents}
\usepackage[parfill]{parskip}

@@

\begin{titlepage}
\hypersetup{pageanchor=false}
\maketitle
\thispagestyle{empty}
\end{titlepage}
\hypersetup{pageanchor=true}
%Below is an example of the problem environment
\chapter{Type Theory}
\section{Introduction}
Type theory is a formal language and deductive system, that is self sufficient in the sense that it need not be formulated as a collection of axioms on top of some other formal system like First Order Logic, instead its deductive system can be built on top of its own formal language. 
\todo{longer sentence to check formating }

@@

\section{Type Construction Operations}
Let's have a look at some important type constructions.
\begin{itemize}
	\item Given types $A, B : \mathcal{U}$ we can define the type $A \rightarrow B$ of functions from $A$ to $B$. 
		We can use $\lambda$-abstraction to construct elements of this type. $\lambda x. \Phi$ lies in $A \rightarrow B$ iff for $a : A$ we have $\Phi[a/x] : B$. 
		For $f : A \rightarrow B$ and $a : A$ we have that the application of $f$ on $a$, denoted as $f(a)$ or $f\;a$, lies in $B$.
	\item Given some type $A : \mathcal{U}$ and a family of types $B$ over $A$, $B : A \rightarrow \mathcal{U}$, we have the type of dependent functions $$\prod_{a : A} B(a)$$
	where for $f : \prod_{a : A} B(a)$ and $x : A$ we have $f(x) : B(x)$. 
	As in the case of non-dependent functions, we can use lambda abstraction to construct elements of a dependently-typed function type.
	\item Given $A, B : \mathcal{U}$ we can define the product type $A \times B : \mathcal{U}$. 
	For $a : A$ and $b : B$ we have the pair $(a,b) : A \times B$. 
	We also have the projection functions 
	$$\mathtt{pr}_1 : A \times B \rightarrow A : (a,b) \mapsto a$$
	$$\mathtt{pr}_2 : A \times B \rightarrow B : (a,b) \mapsto b$$
	\item Given $A : \mathcal{U}$ and family of types $B$ over $A$, $B : A \rightarrow \mathcal{U}$, we can define the dependent pair type $$\sum_{a : A} B(a)$$
	Given $x : A$ and $b : B(x)$ we can construct the pair $(x,b) : \sum_{a : A} B(a)$. 
	We have two projection functions, similar to the case of the product type.
	\item Given $A, B : \mathcal{U}$ we can construct the coproduct type $A + B$. 
	We can construct elements of $A + B$ using the functions
	$$\mathtt{inl} : A \rightarrow A+B$$
	$$\mathtt{inr} : B \rightarrow A+B$$
	This induces the induction principle
	$$\mathtt{ind}_{A+B} : \prod_{C : A+B \rightarrow U} \Big(\prod_{a : A}C(\mathtt{inl}\;a)\Big) \rightarrow \Big(\prod_{b : A}C(\mathtt{inr}\;b)\Big) \rightarrow \prod_{x : A+ B}C(x)$$
	\item Given $x, y : A$ we have the \textbf{identity type} $x =_A y$. 
	An element of this type amounts to a proof that $x$ and $y$ are equal. 
	Say $x$ and $y$ are judgmentally equal. 
	This is captured by the element $\mathtt{idp}_x : x =_A y$. 
	The relevant induction principle describes how we can use elements of an identity type
	$$\mathtt{ind}_{=_{A}} : \prod_{C : \prod_{x,y : A} (x =_A y) \rightarrow \mathcal{U}} \Big( \prod_{x : A}C(x,x,\mathtt{idp}_x) \Big) \rightarrow \prod_{x,y : A} \prod_{(p : x =_A y)} C(x,y,p)$$
	The relevant computation gives us the judgmental equality 
	$$\mathtt{ind}_{=_A} (C,c,x,x,\mathtt{idp_x}) \equiv c\; x$$
	We can concatenate those paths whose domains and codomains allow for it. 
	Paths are equivalences. That is if $p : x = y$ is such a  path, we can provide its inverse $p^{-1}$ for which we have in turn a path between $p\cdot p^{-1}$ and $\mathtt{idp}_y$ and another one between $p^{-1}\cdot p$ and $\mathtt{idp}_x$. 
	\item For every type $A$ there is its \textbf{propositional truncation} $\| A \|$. 
	For every element $a : A$ there exists $|a| : \| A\|$. 
	For every $x, y : \| A\|$ we have $x = y$. 
	Given mere proposition $B$ and $f : A \rightarrow B$, the recursion principle gives us $g : \|A\| \rightarrow B$ such that $g(|a|) \equiv f(a)$ for all $a : A$.

@@

\section{Important types}
\begin{defn}
We call a type $A$ a \textbf{mere proposition} if for every $a,b : A$ we have $a = b$.
\end{defn}

@@

\section{Logic}
Our informal deductions in Type Theory will be reminiscent of First Order Logic ones. 
To be able to use a similar verbiage, we will set down a handful of types, corresponding to the connectives that let us form well-formed formulas in FOL. 
These types need to be mere propositions, so that we can form non-constructive deductions. 
This approach is called `propositions as h-propositions' in the HoTT book. 
In the following, $A$ and $B$ are mere propositions.
\begin{itemize}
\item When we talk of conjunction $A \wedge B$, we mean the product $A \times B$. 
\item We interpret $A \vee B$, as the truncation $\| A + B\|$.
\item We interpret $\forall a\in A,\; P(a)$, where $P(a)$ is a mere proposition for all $a \in A$, as $\prod_{a : A}P(a)$. 
\item We interpret $ \exists a \in A,\; P(a)$, where $P(a)$ is a mere proposition for all $a \in A$, as $\| \Sigma_{a : A} P(a) \|$
\end{itemize}
The use of the above notation is interchangeable in the sections to follow.

\chapter{Modalities}
%Under the view of types as $(\infty, 1)$-categories, functions between them are functors. 
%We can then view the universe $\mathcal{U}$ itslef as a category. 
\begin{defn}
A \textbf{modality} is a function $\bigcirc : \mathcal{U} \rightarrow \mathcal{U}$ with the following properties.
\begin{enumerate}
	\item For every type $A$ we have a function $\eta_A^\bigcirc : A \rightarrow \bigcirc\; A$
	\item for every $A : \mathcal{U}$ and every type family $B : \bigcirc A \rightarrow \mathcal{U}$ we have a function
	$$\mathtt{ind}_\bigcirc : \Big( \prod_{a : A}\bigcirc (B (\eta_A^\bigcirc\; a)) \Big) \rightarrow \prod_{z : \bigcirc A} \bigcirc (B\; z)$$
	\item For every $f : \prod_{a : A} \bigcirc (B(\eta_A^\bigcirc\; a))$ there is a path $\mathtt{ind}_\bigcirc (f)(\eta_A^\bigcirc\; a) = f\; a$
	\item For all $z,z' : \bigcirc \; A$, the function $\eta_{z=z'}^\bigcirc : (z = z') \rightarrow \bigcirc (z = z')$ is an equivalence.

@@

\chapter{LLPO}
\begin{defn} 
The Lesser Limited Principle of Omniscience, states that given binary sequence $s : \mathbb{N} \rightarrow \mathbf{2}$ and the fact that there is at most one occurence of $1$ in the sequence, formally 
$$\mathtt{atMost1one} :\prod_{n_1 : \mathbb{N}} \prod_{n_2 : \mathbb{N}} s(n_1) = 1 \rightarrow s(n_2)= 1 \rightarrow n_1 = n_2$$
we can then have by the LLPO a witness for $\mathtt{p_{odd}} \vee \mathtt{p_{even}}$, where $\mathtt{p_{odd}}$ (with $s$ as an implicit argument) is the statement that for all odd positions $n$, $s(n) = 0$, formally $\mathtt{p_{odd}} \equiv \prod_{n : \mathbb{N}} \mathtt{odd}(n) \rightarrow s(n) = 0$. 
Similarly for $\mathtt{p_{even}}$. 
\end{defn}
LLPO can be viewed as a weaker form of the Law of Excluded Middle.

\begin{lem} 
The Law of Excluded Middle and Markov's Principle implies the Lesser Limited Principle of Omniscience.
\end{lem}

@@

	\item $\neg \mathtt{p_{odd}}$, alongside LEM, implies that there exists odd $n_e : \mathbb{N}$ such that $s(n_e) = 1$. 
	We can now prove $\prod_{n : \mathbb{N}} \mathtt{even}(n) \rightarrow s(n) = 0$. 
	Let $n : \mathbb{N}$ such that $\mathtt{even}(n)$ is true. 
	By the definition of $s$, $s(n) = 0 \vee s(n) = 1$.
	We invoke the principle of induction of coproducts and prove LLPO from the disjuncts. 
	\begin{itemize}
		\item $s(n) = 0$, in this case we are done.
		\item $s(n) = 1$. 
		By $\mathtt{atMost1one}$ we have $n = n_e \Rightarrow $ $n$ even and odd which is a contradiction $c : \bot$. 
		Ex Falso,  $\mathtt{efq} : \bot \rightarrow s(n) = 0$. 
		Then $\mathtt{efq}(c) : s(n) = 0$.
	\end{itemize}
\end{itemize}
\end{proof}

@@

\chapter{Extensional Type Theory}
There exists a model of extensional type theory that validates the following `Independence Principle'
$$\bigg( \prod_{s : \mathbf{2}^\mathbb{N}} P\; s \rightarrow \Big(\prod_{n : \mathbb{N}}s\; n = 0 \Big) \rightarrow \Big\lVert \sum_{z : \mathbb{N}} Q\; s\; z \Big\rVert \bigg)$$
$$\rightarrow \bigg( \prod_{s : \mathbf{2}^\mathbb{N}} P\; s \rightarrow  \Big\lVert \sum_{z : \mathbb{N}} \Big(\prod_{n : \mathbb{N}}s\; n = 0 \Big) \rightarrow Q \; s\; z\Big\rVert \bigg)$$
where $P: \mathbf{2}^\mathbb{N} \rightarrow \mathtt{Prop}$ and $Q : \mathbf{2}^\mathbb{N} \rightarrow \mathbb{N} \rightarrow \mathtt{Prop} $ are families of propositions. 
Putting it plainly, if the left hand side of the above is true, then $z : \mathbb{N}$ does not depend on the proof of the constantness of $s$ to $0$.

@@

In this section we work in the context of this model of Extensional Type Theory.

Let $A :\equiv \sum_{a: \mathbf{2}^\mathbb{N}}\mathtt{atMost1one}\;a$. 
When referring to elements of $A$ we implicitly mean the first part of the pair. 
Let
$$B : A \rightarrow \mathcal{U}$$
$$B :\equiv \lambda\; a.\; \mathtt{p_{odd}}\; a + \mathtt{p_{even}}\; a$$

We would like to have access to the following instance of IP in our intensional model as well.
\begin{multline*}
\prod_{a: A} \prod_{f: B\; a \rightarrow \mathbb{N}} \Bigg( \bigg( \prod_{s : \mathbf{2}^\mathbb{N}} \Big(\big(\prod_{n : \mathbb{N}} s(n) = a(2 \cdot n)\big) \vee \big(\prod_{n : \mathbb{N}} s(n) = a(2\cdot n +1)\big) \Big) \rightarrow \\
	 \Big(\prod_{n : \mathbb{N}}s\; n = 0 \Big) \rightarrow \Big\lVert \sum_{z : \mathbb{N}} \prod_{p : B\; a} f^*\; p = z \Big\rVert \bigg)
\end{multline*}
\begin{multline*}
\rightarrow \bigg( \prod_{s : \mathbf{2}^\mathbb{N}} \Big(\big(\prod_{n : \mathbb{N}} s(n) = a(2 \cdot n)\big) \vee \big(\prod_{n : \mathbb{N}} s(n) = a(2\cdot n +1)\big) \Big) \rightarrow \\
	  \Big\lVert \sum_{z : \mathbb{N}} \Big(\prod_{n : \mathbb{N}}s\; n = 0 \Big) \rightarrow \prod_{p : B\; a} f^*\; p = z \Big\rVert \bigg) \Bigg)
\end{multline*}
The fact that $\mathbb{N}$ is modal, follows from this in Intensional Type Theory. 
Our plan of action shall be to find a consequent of it, IP', strong enough for our purposes, that has the right form so that by Theorem 6.1 of \cite{1905.03014} we have a model of intensional type theory that satisfies IP'.

We drop the truncation around $\lVert \sum_{z : \mathbb{N}} \prod_{p : B\; a} f^*\; p = z \rVert$ and since this is the consequent of the antecedent, the resulting statement is weaker than the original IP. We then uncurry 4 times to reach what we propose as our IP', a function that takes four arguments in the form of a quaternary dependent pair

\begin{align*}
&a: &A
\\
&f: &B\; a \rightarrow \mathbb{N}
\\ &\_ :
\begin{split}
\bigg( \prod_{\bar{s} : \mathbf{2}^\mathbb{N}} \Big(\big(\prod_{n : \mathbb{N}} \bar{s}(n) = a(2 \cdot n)\big) \vee \big(\prod_{n : \mathbb{N}} \bar{s}(n) = a(2\cdot n +1)\big) \Big) \rightarrow \\
	\quad \Big(\prod_{n : \mathbb{N}}\bar{s}\; n = 0 \Big) \rightarrow  \sum_{z : \mathbb{N}} \prod_{p : B\; a} f^*\; p = z  \bigg)
\end{split}
\\
&r : & \sum_{s : \mathbf{2}^\mathbb{N}} \Big(\big(\prod_{n : \mathbb{N}} s(n) = a(2 \cdot n)\big) \vee \big(\prod_{n : \mathbb{N}} s(n) = a(2\cdot n +1)\big)
\end{align*}
and has return type
$$\Big\lVert \sum_{z : \mathbb{N}} \prod_{p : B\; a} f^*\; p = z \Big\rVert$$

\begin{lem}\label{NisBnull}
$\mathbb{N}$ is $\lVert B \rVert$-null.
\end{lem}
\begin{proof}
Given $a : A$ and $f : \lVert B \rVert \; a \rightarrow \mathbb{N}$ we need to prove that there exists unique $f' : \mathbf{1} \rightarrow \mathbb{N}$ through which $f$ factors, in the sense that $f = f' \circ g$, where $g : C\; a \rightarrow \mathbf{1}$ is the sole inhabitant of its function space. 
Functions with $\mathbf{1}$ as their domain are constant. 
We therefore need to find some $z : \mathbb{N}$ so that $\lambda\; \_.\; z$ is $f'$. 
To that end, we will use the following instance of the Independence Principle
\begin{multline} \label{IPlhs}
\bigg( \prod_{s : \mathbf{2}^\mathbb{N}} \Big(\big(\prod_{n : \mathbb{N}} s(n) = a(2 \cdot n)\big) \vee \big(\prod_{n : \mathbb{N}} s(n) = a(2\cdot n +1)\big) \Big) \rightarrow \\
	 \Big(\prod_{n : \mathbb{N}}s\; n = 0 \Big) \rightarrow \Big\lVert \sum_{z : \mathbb{N}} \prod_{p : B\; a} f^*\; p = z \Big\rVert \bigg)
\end{multline}
\begin{multline} \label{IPrhs}
\rightarrow \bigg( \prod_{s : \mathbf{2}^\mathbb{N}} \Big(\big(\prod_{n : \mathbb{N}} s(n) = a(2 \cdot n)\big) \vee \big(\prod_{n : \mathbb{N}} s(n) = a(2\cdot n +1)\big) \Big) \rightarrow \\
	  \Big\lVert \sum_{z : \mathbb{N}} \Big(\prod_{n : \mathbb{N}}s\; n = 0 \Big) \rightarrow \prod_{p : B\; a} f^*\; p = z \Big\rVert \bigg)
\end{multline}
where $f^*$ is the composition $f \circ C^B$, with $C^B$ being the canonical function $B\; a \rightarrow C\; a$. 
Note that the left hand side \ref{IPlhs} is provable. 
So we have a witness of the right hand side \ref{IPrhs}.
By providing the odd $s_1$ and even $s_2$ subsequences of $a$ to it, we get hold of $z_1 : \mathbb{N}$ and $z_2 : \mathbb{N} $ respectively. 
Equality on $\mathbb{N}$ is decidable, therefore $(z_1 = z_2) + (z_1 \neq z_2)$ is provable.

We will first prove that the existence of a candidate for $f'$ follows from both constituents of the coproduct.
\begin{itemize}
\item First the case where $z_1= z_2$. 
We arbitrarily pick $z_1$ and propose $f' : \equiv \lambda \;\_.\; z_1 : \mathbf{1} \rightarrow \mathbb{N}$. 
By function extensionality we reduce proving $f= f' \circ g$ to proving $f\; c = f'\circ g\; c$ for arbitrary $c : C\; a$. \\
Since our goal is a mere proposition, we can act as if we have access to an actual $b : B\; a$. 
We then take cases on $b$.  
\begin{itemize}
\item We first consider the case $\mathtt{p_{odd}}\; a$. 
Through this we can still get access to some actual $b' : B\; a$. 
The odd subsequence is constantly $0$, which we can provide to the rhs of the IP alongside with $b'$ (and $s_1$ as the first argument) to deduce $f^*\; b' = z_1$. 
Using this we then prove $f\; c = z_1$. 
We concatenate this with $f' \circ g \; c = z_1$ to conclude this case. 
\item The $\mathtt{p_{even}}$ case is very similar. 
The only difference is that we provide $s_2$ as the first argument to the rhs of the IP and then have $z_1=z_2$ be part of the concatenation of paths when we prove homotopy between $f$ and $f' \circ g$. 
\end{itemize}
